### 2004-12-01 Article/article_renpar.tex
*** empty log message ***
git-svn-id: https://scm.gforge.inria.fr/svn/cigri/trunk@236 a0e46a8e-710f-0410-a56b-8ab1cdea0c2d

%=======================================================================
% Fichier      : article_renpar.tex
% Title        : 
% Auteur       : Estelle Gabarron <Estelle.Gabarron@imag.fr>
% Creation     : 01 Dec 2004
% Modification : 
%=======================================================================

%--- Chargement de package ---
\documentclass[11pt,twoside,a4paper]{article}

\end{document}