### 2014-09-01 articles/2014/whisper/aricleEbauche/Draft_BriandBzeznikIrodsUserMeeting2014.tex
samll change, tips data management system

%\documentclass[a4paper,twoside]{article}
\documentclass[a4paper, 10pt]{article}
\usepackage{graphicx}
\usepackage{rfia2000}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{multicol}
\usepackage{caption}
\usepackage{listings}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\setlength{\parindent}{0.3cm}

\begin{document}
\long\def\/*#1*/{}
\date{}%{\today}
\title{ ~\\ \Large\bf Data Intensive processing with iRODS and the middleware CiGri for the Whisper project}

\author{\begin{tabular}[t]{c@{\extracolsep{8em}}c}
	Briand Xavier\footnote{Cnrs, Isterre, Whisper, Verce, email xav.briand@gmail.com}  & Bzeznik Bruno\footnote{CIMENT, Universit\'e Joseph Fourier, Technical header of CIMENT} \\
\end{tabular}}
\maketitle
%\thispagestyle{empty}
\subsection*{Abstract}
{\em
Like many projects in science of the universe, the seismological project Whisper is faced with massive data processing.
This leads to specific IT software for the project as well as suitable IT infrastructure. We present here both aspects.

We provide a flexible way to design a sequence of processing. We also deal with data-management and computational optimization questions.
On IT infrastructure, we present the platform CIMENT provided by the University of Grenoble. It offers a data-grid environement with the 
distributed storage iRODS and the grid manager CiGri.

This is the partnership between these two IT components that has enabled a data-intensive processing and, also,  permits to the Whisper project to bring new scientific results.
}
~\\ \\ 
{\bf Keywords}:  Data-Intensive, Grid computing, Distributed storage, Seismic Noise, Whisper, CiGri, iRODS.

\begin{multicols}{2}
\/*
\section{Plan}
\begin{enumerate}
	\item Introduction, abstract (3\S , 1 page) 
  	\item Background (7-10\S, 1-2 pages)
    \begin{enumerate}
  	  \item The Whisper project
  	  \item Size of observational data
  	  \item Size of computational data
  	  \item General Big data/ science of  universe
  	  \item Coupling scientific objectives / IT constraints
  	  \item Data grid paradigm
  	  \item Data grid in Grenoble
  	  %\item Whisper organisation
  	  \item Collaboration Whisper/IT infrastructure
  	\end{enumerate}
  	\item Software for data-intensive processing (6-12\S , 1-2 pages)
  	\begin{enumerate}
      \item General IT codes whisper
      \item Computer language and librairies
      \item Structure of codes
      \item Package for raw data
      \item Package for correlations
      \item First optimisation of correlation
      \item Second optimisation of correlation
      \item Codes for analysis of correlations
      \item Where we can run the codes
    \end{enumerate}
  \item IT infrastrcuture for grid computing (6-12\S , 1-2 pages)
    \begin{enumerate}
  	  \item Needs of coupling storage computation
  	  \item Presentation CIMENT
  	  \item Coupling iRODS/CiGri
  	  \item Resulting Data-Grid
      \item Presentation iRODS
	  \item General Infra iRODS (diagram)
	  \item Effective Nodes iRODS (diagram)
  	  \item General CiGri
  	  \item Mechanism with OAR
  	  \item Mechanism resubmission/besteffort
  	  \item Description of a campaign
  	  \item Low Interaction CiGri/iRODS
  	  \item CiGri V2 and V3 functionalities
  	\end{enumerate}
  \item Results and feedback (28-31\S, 6-7 pages))
      \begin{enumerate}
	    \item Whisper Use Case
        \item Infrastructure CIMENT experiences
       \end{enumerate}
\end{enumerate}
*/

\newpage
\section{Introduction}% 7-10\S, 1-2 pages}

The \emph{Whisper} \footnote{FP7 ERC Advanced grant 227507, see whisper.obs.ujf-grenoble.fr} 
project is a an european project on seismology whose goal is to study properties of the earth
with the seismic ambient noise such that evolution of seismic waves speed. 
This noise is almost all the signal continuously recorded by
the seismic stations worldwide (Europe, China, USA, Japan), except earthquakes. It offers new
observables for the seismologists, new types of virtuals seismograms that are not only located at the 
place of earthquakes and that are provided by the operation of correlation  which requires significant computations.
For instance, one can obtain wave paths that probes the deepest part of the Earth \cite{key:BPCPBR, key:PCP}.
%(variations in space and time of the seismic waves)

Accordingly, this is one of the first project in the seismological community that studies
systematically the continuous recordings, which represents a large amount of seismological data, 
of the order of several tens of terabytes.
For instance, one year of the Japanese Network is about 20 TB or 3 months of the mobile network
USArray represents 500 GB (it depends on the sampling of the recorders).

In addition, the calculation operations downstream may produce even more data than the observation data.
To give an order of magnitude, more than 200 TB have to be processed by the Whisper project at the same time. 
A classical processing produces 8 TB in 5 days. Another computation 'read' 3 or 4 TB and 'produced' 1 TB in 6 hours.
Many tests of the signal processing are done and computational data are deleted as and when required.
~\\ \\

\vspace{0.18cm}
Nowadays, the earth sciences or more generally, sciences of the universe are widely engaged in
data-intensive processing. This leads to design scientific workflow, towards data-intensive discovery and e-Science.
\vspace{-0.04cm}

Reflected by the Whisper project, we have to organize the science objectives with
the computer constraints. We have to take into account 
the duration of postdocs and PhD theses, as well as the availability of computer infrastructures and their ease of access. 
This leads to many questions about software development, including the genericity of computer code and 
the technical support. But it also influences in terms of choice of appropriate infrastrutcures.
%\vspace{-0.02cm}

Even if this project has his own ressources, such a problem of data-intensive requires specific tools able to 
organize distributed data management and acces to computational ressources: a data grid environment.
\vspace{-0.04cm}

The University of Grenoble offers, thanks to the High Performance Computing (HPC) centre \emph{CIMENT}, this kind of environment with the data management system \emph{iRODS}
and the middelware \emph{CiGri}.% (CIMENT platform).
\vspace{-0.04cm}

It is thanks to the close collaboration between IT ressources of Whisper and the infrastructures of the University
that this project has been implemented as we show below.

~\\ \\

%\newpage

\section{Software for data-intensive processing} \label{Section_Software_for_data-intensive_processing} %(6-12\S , 1-2 pages) }

A part of the Whisper project is specificaly dedicated to the IT codes. This includes to design a specification, an implementation and some optimisations of the sequence of data-management as well as of the computations \footnote{see code-whisper.isterre.fr/html/ (part of the design)}.
This project uses its own IT resources (servers, dedicated bay) as well as the common IT infrastructure of the university.
We developed also adaptations for the IT infrastructures and we provide technical support for researchers.

Most of the IT codes are written with the \emph{Python} language and make intensive use of the scientific libraries \emph{Scipy} (fortran and C embeded) 
and \emph{Obspy} \footnote{ see www.obspy.org} (essentially the 'read' function) which is dedicated to the seismological community.

The IT codes consist of several tools described schematically in figure \ref{whisperGeneralWorflow} and grouped into three parts. The first one concerns the signal processing, the second part permits the computation of the correlations and the last part consists of codes for the analysis of the correlations.

The first package provides a flexible way to process raw data, to specify a pipeline of pre-processing of signal.
The user starts by specifying a directory, a set of seismic stations and a set of dates. Then the codes scans the directory
and extracts all pieces of seismograms (also called traces) and rearranges them in a specific architecture of files in order to calculate the correlations to the next step.
We use here intensively the function 'read' of the library Obspy which allows to open most of the seismogram file formats.
The user also define his own sequence of processings. He can use the functions predefined but also the Python libraries he needs and,
moreover, he can add eventually his own functions.

The second package concerns correlations. 
Roughly speaking, a correlation is an operation with two seismograms (for a given time window) that provides the coherent part of the two seismograms 
(associated to the 2 stations) which is the seismic waves that propagate between the 2 stations.
(Moreover, in some favorables cases, it converges to the Green's function). 
Thus, the code computes all the correlations and provides an architecture of files that corresponds to all the couples of seismograms (for each date).

%\end{multicols}
\begin{center}
\centering
\captionof{figure}[h]{\label{whisperCorrelationStep} Step of the correlations}
\includegraphics[width=6cm]{schemaCorrelationStep.png}
\end{center}
%\begin{multicols}{2}

Note that the space complexity is linear for seismograms processing but quadratic for the correlations. We have therefore to store the seismograms processing before compute the correlation in order to benefit of the good complexity.
These quadratic space complexity can be critical and lot of effort was made in order to optimize the computation in two direction. First we improve the computation of the fast fourier transform by pre-calculating some "good" combinations of small primes numbers. 
With this method, we improve of forty percent the time computation in the favorable cases.

Nevertheless, the main optimization was made by testing the behaviour of the carbage collector of Python in order to follow the cache heuristics. More precisely, we do not use the 'gc' module or the 'del' statement but we try to schedule and localize the line of code in order to find the good unfolding that uses the architecture optimally.

The last part of computer codes concerns the analysis of correlations (the virtual new seismograms) with methods such as beamforming, doublet or inversion.
We also compute correlations of correlations C3 (also new seismograms). For example, we study the variations
in velocity of seismic waves as we illustrate below in figure \ref{JapanVelocity}.

These codes permit to process a dataset on a computer laptop. 
Nevertheless, to take advantage of IT infrastructure at the University of Grenoble, 
adjustments have been made for the grid computing as we shall see later.

\end{multicols}
\begin{figure}[h]
\centering
\caption{\label{whisperGeneralWorflow} Main sequences of processings of the Whisper Codes}
\includegraphics[width=14cm]{briandPosterAGU2013GimpProcessing.png}
\end{figure}

\begin{multicols}{2}

\section{IT infrastructure for grid computing}% (6-12\S , 1-2 pages) }

\subsection{CiGri and iRODS synergy}

The data-intensive processing
needs obviously an IT infrastructure in order to couple storage and computation.
In our cases, most of the processings are embarrassingly parallel.
The amount of data and the location of available compute nodes suggests the use of a 
distributed storage system and a grid manager.

The IT infrastructure used here is provided by the \emph{CIMENT} \footnote{see CIMENT.ujf-grenoble.fr} 
platform. CIMENT is the High Performance Computing (HPC) center of the Grenoble University. It offers a partial pooling of computing 
resources (10 computing clusters, 6600 cores and some GPUS) and many documentations for users. 
%It also provides a data grid environement : a grid of supercomputers and a distributed storage.
Moreover, the computational resources are integrated in a local grid of supercomputers. Associated with a distributed storage,
it provides a local data grid environement.

The distributed storage accessible by all the computing nodes of all the clusters is managed by iRODS \footnote{see irods.org}.
Nowadays it represents approximately 700 TB. The grid computing is managed by the \emph{CiGri 
\footnote{see CIMENT.ujf-grenoble.fr/cigri/dokuwiki} middleware}, that is part of the OAR \footnote{see oar.imag.fr} \cite{key:CDGHMMNR, key:CHR2009} project (the Resource and Job Management System on which CiGri relies). CiGri and iRODS together build a complementary solution for embarrassingly parallel computations with large input/output distributed data sets.

Furthermore, whith unitary parametric jobs that are reasonnably short in time, CiGri can deal with the best-effort mode provided by OAR. In this mode, grid jobs are scheduled on free resources with a zero priority and may be killed at any time when the local demand of resources increases. This CIMENT organization (independant computing clusters glued together with a best-effort grid middleware and a distributed storage), in place for more than a decade, has proven to be very efficient, allowing near one hundred percent usage of computing resources thanks to small jobs being managed at the grid level.

Furthermore, as the results of the grid jobs are stored into the distributed storage with a unique namespace, iRODS also acts to the user as a centralized controller with a 
total observation and thus allows the user to monitor its calculation.
~\\

\subsection{iRODS infrastructure}

The Integrated Rule-Oriented Data System (iRODS) is a data managment system offering a single namespace for files that are stored on differents resources that may be on different locations. The administrator can set up rules (microservices) to perform some automatic actions, for example the storage of a checksum or an automatic replication to the nearest resource (staging).  The user can control himself replications and create user-defined metadata. iRODS exposes a Command Line Interface (the i-commands), an API useable from several programming languages (C, python, PHP,...), a fuse interface, a web gui, and a webdav interface. The meta-catalog is an SQL database, which makes it very efficient for managing additionnal meta-data or making advanced queries (see \cite{key:CCQSC2005} for an illustration of use). It is not "block-oriented", and thus relies on underlying Posix filesystems. Performance is not the main goal, but when the files are distributed on different resources, the only bottleneck is the meta-catalog (which is centralized).
\\

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaCimentStorageNodesIcat} iCAT and storage nodes}
\includegraphics[width=8cm]{schemaCimentStorageNodesIcat.png}
\end{center}

The iRODS storage infrastructure of CIMENT consists of a unique zone with the iCat server and a dozen of nodes as illustrated on figure \ref{schemaCimentStorageNodesIcat}. The nodes are groupped inside 3 different locations, called site A, site B and site C, having heterogeneous WAN connexions. Each site has it's own 10Gbe local network switch.

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaCimentStorageSupercomputers} iRODS resources close to supercomputers}
\includegraphics[width=8cm]{schemaCimentStorageSupercomputers.png}
\end{center}
Those 3 sites are located into the 3 main datacenters where the CIMENT supercomputers live, so there are always close storage resources with the computing nodes (Figure \ref{schemaCimentStorageSupercomputers}). All resources of a given site are groupped into an iRODS resourceGroup and a rule tells that if a data is to be written from a computer of this site, then the data is written by default on a resource randomly chosen inside this group. So, data are always written to a local iRODS resource, using the LAN and not the WAN.
Note that site C has only 1Gbe WAN connexions while A and B have 10Gbe WAN connexions (Figure \ref{schemaCimentStorageNodesIcat}). So, to optimize, we've set up automatic staging for site C: when a data is get from site C and the meta-catalog tells that the file is located on a site A or site B resource, then the file is automatically replicated to a resource of site C so that if it is accessed again later, it is not more transfered through the 1Gbe WAN link.

Capacity has now reached 700 TB and is constantly evolving and increases with investment in new projects, as iRODS offers a great scalability by simply adding new storage resources.
Each node has currently 2 RAID arrays from 24 to 48 raw TB as illustrated at the figure \ref{schemaNodeStorageCiment}

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaNodeStorageCiment} A storage node}
\includegraphics[width=5cm]{schemaNodeStorageCiment.png}
\end{center}
iRODS nodes are running Debian GNU/Linux with Kanif \footnote{see http://taktuk.gforge.inria.fr/kanif/} \cite{key:CHR2009} for easy synchronisation of the system administration.
CIMENT has set up a web interface where the user can easily check the status of the resources (figure \ref{schemaCimentStorageStatus}).

\begin{center}%[h]
\centering
\captionof{figure}{\label{schemaCimentStorageStatus}}
\includegraphics[width=8cm]{schemaCimentStorageStatus.png}
\end{center}

\subsection{CiGri infrastructure}

The access to 6600 cores of the clusters of the CIMENT platform is achieved through the middleware CiGri.
CiGri launches embarrassingly parallel jobs on idle processors of every computing clusters and then optimizes the resources usage which are used for parallel jobs otherwise.

Each cluster of the University of Grenoble uses the resource manager OAR. CiGri acts, among other things, as a metascheduler of OAR.
It retrieves the clusters states trough OAR and submits the jobs on free resources without exhausting the local scheduling queues.

%Figure serveur CiGri et clusters, REST API

While it may work in normal mode, CiGri is mostly used in best-effort mode and thus provides an automatic resubmission mecanism. CiGri offers a customizable notification system with a smart events management. With those mecanisms, the user can submit a big amount of small jobs, called a campaign, and forget about it until all the jobs of the campaign are terminated or CiGri notifies a serious problem.

Roughly speaking, in order to run a campaign, the user describes through a file (in the JSON format) the parameters of the campaign 
such as the accepted cluters, the needed resources, the maximum duration, the location of the codes, a prologue or epilogue script,...  
Codes and input data are retrieved from iRODS using i-commands into the prologue scripts and the jobs scripts (or using the iRODS API if the jobs are written into a supported language). So, there's no direct connexion between CiGri and iRODS, but the usage is totally complementary through the jobs scripts. 
Moreover, the user defines also a file where each line represents a value of the parameter for the user's code.
Thus, the number of line of these parameter file corresponds to the number of jobs of the campaign.

%Figure JDL+Params-> campaign

Users may monitor their campaigns and acts on it whith a CLI or a REST API. Some statistics are provided, such as the completion percentage (in the term of number of terminated jobs), jobs execution rates, automatic re-submissions rate. When a job fails with a non-zero exit status, the user is notified by mail or jabber and requested for an action before submitting further jobs on the same supercomputer: simple acknowledge, acknowledge and re-submission or abort the campaign. Standard and error outputs of the jobs may be easily retrieved from the CiGri host without having to log on the underlying supercomputers. Users may even not be authorized to log-on to a specific supercomputer but allowed to start and manage best-effort jobs on it thanks to CiGri.

CiGri is now at the version 3, which represents a major evolution in terms of modeling and technology (Rest API, Ruby). It is structured around a PostgreSQL database and high level components (Ruby scripting language, Apache with SSL authentication, Sinatra,...).

\subsection{Authentication and security}

CIMENT has a centralized LDAP infrastructure. Users have the same login on all supercomputers and on the CiGri frontend. As iRODS does not offers a simple and direct LDAP authentication mechanism, we use the simple password method with an automatic synchronisation from our LDAP server to the iRODS database. We also have a script that automatically initializes the iRODS unix environment directly into the users home directory (.irods/.irodsEnv and .irods/.irodsA files) on every supercomputer, so that iRODS authentication becomes completely transparent to the users.

Each site has a filtering router acting as a firewall. As we want all communication schemes to be possible between each iRODS resource (a file might be transfered from a resource to another regardless of the site), we had to open some tcp and udp ports on those firewalls. The range of ports may be defined by the administrator into the iRODS servers configuration file, so that's not an issue.

\section{Results and feedback} % (28-31\S, 6-7 pages)}

\subsection{Whisper Use Case}
\/*
	\begin{enumerate}
	  \item Scientific whisper results: theses, articles, prices
  	  \item Organisation with a DataSet: Evaluation local/CIMENT, capacities adaptation students
  	  \item Comparison data management, examples Japan processing/USArray 
  	  \item Evolution specification, genericty of the codes, degree of automatization
  	  \item First step: Conversion data format, distributions of datas on resources: decrease the concurrency
  	  \item Back to the first package signal processing for Japan
  	  \item Identification seismogram: Dates, stations  / Modelization distribution: sublist Dates, Stations
  	  \item Adding python modules for data retrieve + Calls.
  	  \item Focus on encapsulation queries
      \item The file parameters
      \item the file start.bash
      \item The jdl file
  	  \item Back to the package correlation: similar adds 
  	  \item Focus on the distribution problem (Space complexity/Transfer)
  	  \item Focus on file size: flat/dictionnary
  	  \item Mode besteffort: increase of datas transfert, it possibly adds also new steps of data storage.
  	  \item Some order of magnitude for the Japan Computation.
  	  \item On the Japanese results.
  	\end{enumerate}
*/

Whisper is one of the projects that have made it possible to ensure that the seismic noise brought new observable.
This permits to carry out several scientific results including imaging and monitoring. 
Concerning monitoring, further studies provide news results about sligth variations of seismics waves induced by earthquakes.
Many articles are in part due to this project as well as
several post-docs and PhD (see \cite{key:CFLC2010, key:BPCPBR, key:HCM2014, key:OPLC2013, key:PCP, key:MPPVCS, key:RCRZCSKCCWD, key:RBCSP, key:BCTASBEM} and also whisper.obs.ujf-grenoble.fr, rubric publication).

Most of the time, on the computer part, the approach with researchers is as follows.
After retrieving data from a data center or directly between persons, we have to assess how this data can be processed.
According to the computing time and storage capacity, either we perform operations on a dedicated bay (also host by CIMENT), or either we use the CIMENT infrastructure.
It depends also on the ease of computer users and most of the time, at least the last treatments (less computationally expensive) are made locally
(Some IT codes are also provided in order to retreive results on distributed storage).
For instance, with small datasets (datas from La Reunion or the Alpes) we work only locally. With larger dataset (China, USArray, Japan Network).
We use both local and distributed computation.

We focus now on the part of processing that use the data grid environment of CIMENT.
But note before that an other aspect, and not least, is the evolution of the specification. 
Often, students and researchers have new requirements and the IT codes evolves with these specification.
We try to be as generic as possible in order to, among other things, to achieve a sufficient level of automatization.
However, sometimes we need to develop some parts specifically because of lack of time.
These IT problems of specification and development time are among the most complex to evaluate for this type of project.

A first step, for the IT part of the scientific workflow, begins by storing data and also by checking their integrity.
We also require that data be in a seismic standard formats.
If conversion is necessary, it can be a large data intensive computing and specific codes are developed.
To minimize concurrency, we have also to ensure that the data are well distributed on iRODS. 
Indeed, it happened that the data are too centralized on a resource, so that is truly diminished processing capacity.
Some python codes are dedicated to this task and can replicate or spread randomly a Collection from a set of resources to another set.

Let us come back to the package of signal processing of Whisper (the first part of the figure \ref{whisperGeneralWorflow}).
Because each seismogram can be treated separatly it is the simplest case for data grid process. We treated for instance one year of the
Japanese seismic Network (HiNet, Tiltmeter and FNet) around of the giant 2011 Tohoku-oki earthquake (6 months before and 6 months after).
Note that first we need to convert 9 Terabytes of Japanese Data into around 20 terabytes of a standard format (here mseed or sac).
Then we try many processing for the 20 TB (filter, whitening, clipping, ...) and store the results in seismogram with a duration of one day.

As almost all seismological data, the Japanese data are identified by the dates, names of seismic station and sub-networks. Therefore
the choice of the modelization, in order to retrieve and distribute the data, follows these seismological metadatas.
More precisely, the modelization of the distribution of the computation is made by setting a subset of dates and a subset of stations 
(This corresponds also to a normal use for researchers that wants to test some processing rapidly with a subset of seismograms).
Note also that the distribution is constant with respect to transfer.

In order to get a set of seismograms from the iRODS storage  to a computational node, we add Python modules to the Whisper package of seismogram processing.
These modules contains classes that permits to retrieve a subset of seismograms (defined by the two subsets described above). More precisely, a first step is either to test directly existence of data or either building of hash tables in order to know the available seismograms.
Then we provide an iterator (in this case, a generator for Python) on available data in order to use other methods that performs the i-command
'iget'. The same approach is made for the storage of the results (with the 'iput').

Schematically, the 'Main' Python module of IT code looks like:

\begin{lstlisting}
...
get parameter
...
build the iget commands
perform iget (encapsulation)
...
Codes whisper 
for seismogram processing 
....
build the iput commands
perform iput (encapsulation)
...

\end{lstlisting}

Note that these commands of transfer between iRODS and computational nodes can become very difficult to achieve because of the concurrency of the queries.
To take into account this obstacle, one develops a module that provides lot of encapsulations of the i-commands (number of try, waiting time,
resubmission, with error, 'else' command,  etc...). The IT infrastructure provide also wery usefull encapsulations.

We have also to adapt our process for the grid computation with CiGri. We first define a file of parameters 'param.txt' where
each line correponds to the parameters of one job on the grid, for instance:

\begin{lstlisting}%[language=bash]

cat param.txt

traces160_0_8_0 160 0 8 0
traces160_0_8_1 160 0 8 1
...
traces160_1_8_0 160 1 8 0
traces160_1_8_1 160 1 8 1
...
\end{lstlisting}

Here we divide the dates in 160 sublists and the stations in 8 sublists. 
The line "traces160\_0\_8\_1 160 0 8 1" correponds to a job named 'traces160\_0\_8\_1' where
we take the sublist of dates of index 0 and the sublist of stations of index 1 (each sublist have the same length +/-1).
(There are also other parameters for components and networks not described here.)

For each of the 1280 jobs, a script is launched by CiGri , say 'start.bash', that take for
arguments a line of the file parameter. The script, among other things, load necessary library like appropriate python,
here 'main.py' and run the code (This is a diagrammatic view).

\begin{lstlisting}%[language=bash]

cat start.bash

#!/bin/bash
set -e
...
module load python
...
NumberDate=$2
IndexDate=$3
NumberStation=$4
IndexStation=$5
...
cd DirToCompute/Codes
python main.py NumberDate 
     IndexDate NumberStation IndexStation
...
cd
rm -rf DirToCompute
\end{lstlisting}

Then one defines the json jdl file (job description language), say 'processingSeismogram.jdl'.

\begin{lstlisting}%[language=bash]

cat processingSeismogram.jdl

{
"name": "test_processing",
"resources": "core=1",
"exec_file": "$HOME/start.bash",
"param_file": "param.txt",
"type": "best-effort",

"clusters": {

"c1": {
  "prologue": [
  secure_iget -f /iRODSColl/start.bash,
  secure_iget -f /iRODSColl/param.txt,
  mkdir -p DirToCompute,
  secure_iget -rf Codes DirToCompute,
  ... other lines of commands ],
  "project": "whisper",
  "walltime": "00:20:00"
  },

"c2": {
  "prologue": [
  ... lines of commands ],
  "project": "whisper",
  "max_jobs": "450",
  "walltime": "00:30:00"
  },
...
}
...
}
\end{lstlisting}

Here the campaign named 'test\_processing' take only one core. It run the file 'start.bash' with the parameter
file 'param.txt' in mode besteffort. It uses the clusters named 'c1' and 'c2' for the project 'whisper' with the duration
defined by 'walltime'. In the prologue, we retrieve script and the parameter file on iRODS as well as the code 'Codes' on iRODS. 
One can add other parameters for each clusters such that the maximum number of
jobs runing at the same time (the variable 'max\_jobs').

Concerning the package for the correlations (the second part of the figure \ref{whisperGeneralWorflow}), we also add similar modules
to retrieve on iRODS the subsets of seismograms that have been treated. In order to compute all the correlations that corresponds to
all the couple of seismograms we have two types of processes. For a fix subset of dates, either we retrieve one subset of stations 
and compute all the existing couples for this subset, or either we retrieve two disjoint subsets of stations and compute all the 
correlations where each seismogram is in a different sublist. This distribution for the computation of the correlations offers a good granularity.

With this kind of distribution for the stations, the transfer bewteen iRODS and the computational node becomes proportional to 
the distribution, i.e. the number of sublists of stations. Therefore we have to maximize the distribution of the dates and minimize it
for the stations in order to obtain a reasonable walltime. Note that it is possible to improve the transfert.
However, a fairly simple improvement have to effect that the distribution become dependent on the distributed architecture of computation.
In order to keep genericity of the codes we do not change this aspect.

Moreover, special attention was given to selecting the size of files to transfer with iRODS.
We do not store each correlation separately, we group them into dictionary to achieve file sizes between 100MB and 500MB most of the time. 
This order of magnitude seems appropriate for the infrastructure iRODS.
For simplicity, we build the grouping according to the parameters of the distribution. This allows to improve the transfer performance.
However, unlike a flat architecture, this forced us to develop codes to retrieve the data. Analysis of some subset of correlations (the 3rd step of the figure \ref{whisperGeneralWorflow}) may request transfer much more than desired.

The mode besteffort also increases the transfert of data because some job are killed and are submitted in an other place.
Moreover in some cases, such that a big walltime, we add new steps of iRODS storage for the process in order to store certain intermediate calculation
(It is not the case for the correlations because of the good granularity). Note that this may represent a significant development effort.

We focus now on the Japanese computation. With the optimizations described in section \ref{Section_Software_for_data-intensive_processing} 
and the data-grid environment, the computation of 350 millions individual correlations of the Japanese Network (especially the dense Hi-Net Network). 
are done at most a few days. This is a big change that can test many treatments in order to find information in the noise.
More precisely the seismogram processing of the Japanese network take half day here (depend obiously on the resampling).
Depending on the types of the correlations (stacking, overlap) and also of the availability of the grid (we suppose an usual case here), 
the computation of all the correlations take between 9h and 20h. The 'iget' i-commands corresponds approximatively to 11TB (with the best-effort mode,
some transfers are carried out several times) whereas the 'iput' i-command corresponds here to 3.3TB.

We illustrate with the figure \ref{JapanVelocity}, one of the analysis of the correlations that represents an image
of change of the velocity of seismic waves in Japan (the giant 2011 Tohoku earthquake, see \cite{key:BCTASBEM}).
\begin{center}%[h]
\centering
\captionof{figure}{\label{JapanVelocity}}
\includegraphics[width=8cm]{JapanVelocity.png}
\end{center}

\subsection{Infrastructure CIMENT experiences}

With the Whisper project, as with some few other projects, we had the opportunity to test and improve the CIMENT IT infrastructure for a data-intensive case. 
Note also that the data-grid environment is used by many other scientifics projects such that research in Particle Physics \cite{key:BB} or genomics research \cite{key:MLBDJ}.
For this kind of project, data management and input/output data flows are a big part of the process regarding the actual computing time for the analysis. With a grid of several thousands cpu-cores, such a project may act as a real distributed attack against our storage infrastructures! So, as a first consequence of the deployment of such computing jobs, we had to implement new "limiting" functionnalities into our infrastructure softwares or configurations. For example, CiGri is able to limit the number of submitted jobs for a given campaign on a given cluster if we know that the jobs concurrency can overload our iRODS infrastructure. As another example, iRODS may be configured to limit the number of simultaneous connections. But in this case, the i-command returns with an error and the job may be aborted. There's a "--retry" optionnal argument, but with a high level of concurrency, it increases the load and may completely exhaust the servers with a ton of retry queries. We then developped a wrapper for the i-commands, called secure-i-commands that implements a retry on some specific error cases  with an incremental delay to prevent from flooding the iRODS servers. Of course, this can be improved as this leads to a waste of computing resources because they are reserved and not used while waiting. But this is at least a security improvement in case the highler level decision processes does not work as expected because of their high complexity.

Regarding the iRODS choice, it was made on the features: we need a distributed file system suitable for a distributed and heterogeneous grid infrastructure composed of several supercomputers having gateways to reach each-other; and we need a unique namespace. We also need the sites to be independent of the others. It means that a site can be out of service, the data available in another site should still be available. The only case when the entire infrastructure is broken is when the centralized meta-catalog is not reachable. But this part of the infrastructure is located in a highly available datacenter and we may also implement a HA meta-catalog. As it is not posix compliant, iRODS allows more control for the admins and the users. For example, replication can be completely controlled. In our case, we do an automatic replication (staging) of the files when they are got from site A or B to site C, for network optimization. But the user may also want to replicate on every sites to get better performances with big jobs campaigns spread on every clusters of the grid. 
Also better than a posix filesystem, the users can add custom meta-data to the files and collections allowing them to retrieve data by making some queries with an SQL-like syntax.

Another interesting aspect of iRODS for a project like Whisper is that we can register into iRODS data that are stored on a dedicated posix filesystem. For example, the Japanese raw data were stored on external storage disks that we copied through USB directly onto a server dedicated to the project. We then added this server as an iRODS resource and then registered the data to make them available to the whole grid. In this example, we also used the access rules to set up a read/write policy for the only concerned users.
\\
\textbf{Open problems and future works}

\begin{itemize}
  \item iRODS has a different transfert policy for "small" files (under 32MB). In this case, files are transiting through the server used by the client instead of a direct connection between the original resource and the client. In a high load context, with a lot of small file transfers, it leads to an overload of the concerned server. A possible solution to that problem could be to share the load among the servers of a given site. For that, we can imagine using a virtual ip address making a round robind on all the ip addresses of the servers of the site.
  \item We issued some network overloads: automatic staging, as set up for a site that has a lower bandwidth than the others may result in a lot of background iRODS processes doing the replication even when connection rate limit is set up on the servers and even when iget commands are aborted. Under some circumstances, it may completely overload a 1Gb/s ethernet link between 2 sites. Another overload encountered was when the user activates the -Q (use RBUDP (datagram) protocol for the data transfer) option. Even from a single node, this option can cause dramatic overload of network interfaces. Maybe iRODS should implement better rate limit control. 
  \item We made some preliminary tests of the python bindings of the iRODS API (called Pyrods). We noticed that when you have a lot of small operations to do from a python code (like in the Whisper project), either creating a lot of small files, or meta-data, Pyrods may be 10 times faster than a shell loop with i-commands inside. We are now making code examples for the other users to be able to take benefits from this method that would have been useful for the Whisper project.
\end{itemize}

\section{Conclusion}

Despite the problems that have been solved or not, we are currently able to process terabytes of data within a few hours.

Note also that the local data grid (iRODS+CiGri) and more generally the platform of CIMENT as well as the OAR tools permit for the Whisper project  to produce significant new results for seismological community in a reduced delay.

Lot of improvements that involve automatization of the processing, and also concerning the scientific workflow have constantly been made.
We may say that Whisper+CiGri+iRODS is a great success!

\section{Acknowledgements}
We thanks Emmanuel Chaljub, the OAR team and the CIMENT team as well as the Whisper project and the Verce project for all their very useful support.

\begin{thebibliography}{9}

\bibitem{key:BB}
C. Biscarat, B. Bzeznik.
Synergy between the Ciment tier-2 HPC centre and the HEP community at LPSC in Grenoble (France)
{\em J. Phys.: Conf. Ser. 513 032008, 2014}

\bibitem{key:BPCPBR}
P Bou\'e, P Poli, M Campillo, H Pedersen, X Briand, P Roux.
Teleseismic correlations of ambient seismic noise for deep global imaging of the Earth.
{\em Geophysical Journal International 194 (2), 844-848.}

\bibitem{key:BCTASBEM}
Brenguier, F., M. Campillo, T. Takeda, Y. Aoki, N. M. Shapiro, X. Briand, K. Emoto, H. Miyake (2014).
Mapping pressurized volcanic fluids from induced crustal seismic velocity drops.
{\em Science, 345, 80-82, 2014.}

\bibitem{key:CDGHMMNR}
Nicolas Capit, Georges Da Costa, Yiannis Georgiou, Guillaume Huard, Cyrille Martin, Gr\'egory Mouni\'e, Pierre Neyron, Olivier Richard,
A batch scheduler with high level components.
{\em Cluster computing and Grid 2005 (CCGrid05), Cardiff, Royaume-Uni, 2005.}

\bibitem{key:CFLC2010}
JH Chen, B Froment, QY Liu, M Campillo.
Distribution of seismic wave speed changes associated with the 12 May 2008 Mw 7.9 Wenchuan earthquake.
{\em Geophysical Research Letters 37 (18).}

\bibitem{key:CCQSC2005}
Gen-Tao Chiang, Peter Clapham, Guoying Qi, Kevin Sale and Guy Coates,
Implementing a genomic data management system using iRODS in the Wellcome Trust Sanger Institute.
{\em BMC Bioinformatics,12(1):361+, September 2011.}

\bibitem{key:CHR2009}
Beno\^it Claudel, Guillaume Huard, Olivier Richard.
TakTuk, Adaptive Deployment of Remote Executions.
{\em Proceedings of the International Symposium on High Performance Distributed Computing (HPDC) (2009) 91-100.}

\bibitem{key:HCM2014}
G. Hillers, M. Campillo, K.-F. Ma.
Seismic velocity variations at TCDP are controlled by MJO driven precipitation pattern and high fluid discharge properties.
{\em Earth and Planetary Science Letters, Volume 391, 1 April 2014, Pages 121-127.}

\bibitem{key:MPPVCS}
Macquet M., Paul A., Pedersen H., Villase~nor A., Chevrot S., Sylvander M. and the PYROPE working group.
Ambient noise tomography of the Pyrenees and surrounding regions : inversion for a 3-D Vs model in a very heterogeneous crust.
{\em accepted in Geophysical Journal International.}

\bibitem{key:MLBDJ}
M\"unkem\"uller T., Lavergne S., Bzeznik B., Dray S., Jombart T.,
Schiffers K., Thuiller W. 
How to measure and test phylogenetic signal. Methods in Ecology and Evolution, 3: 743-756, 2012.

\bibitem{key:OPLC2013}
Obermann, A., T. Plan\`es, E. Larose, and M. Campillo.
Imaging preeruptive and coeruptive structural and mechanical changes of a volcano with ambient seismic noise.
{\em J. Geophys. Res. Solid Earth, 118, 6285-6294. }

\bibitem{key:PCP}
P. Poli, M. Campillo H. A. Pedersen,
Body-wave imaging of Earth's mantle discontinuities from ambient seismic noise.
{\em Science 338 (6110), 1063-1065.}

\bibitem{key:RCRZCSKCCWD}
D. Rivet, M. Campillo, M. Radiguet, D. Zigone, V. Cruz-Atienza, N. M. Shapiro, V. Kostoglodov, N. Cotte,
G. Cougoulat, A. Walpersdorf and E. Daub
Seismic velocity changes, strain rate and non-volcanic tremors during the 2009-2010 slow slip event in Guerrero, Mexico. 
{\em Geophys. J. Int. 2014;196:447-460. }

\bibitem{key:RBCSP}
Rivet D., Brenguier F., Clarke D., Shapiro N.M., Peltier A. 
Long-term dynamics of Piton de la Fournaise volcano from 13 years of seismic velocity changes measurements and GPS observation.
{\em J. of Geophys. Res. 2014; submitted.}

\end{thebibliography}

\end{multicols}
\end{document}